\documentclass{report}

\usepackage{mathpartir}
\usepackage{syntax}
\usepackage{hyperref}

\begin{document}

\title{Foundations and Limitations of Jacinda}
\author {Vanessa McHale}
\maketitle

\tableofcontents

\section{Introduction}

Jacinda is an expression-oriented, functional language that performs a subset of Awk's function.

In many ways it is unsatisfactory or even defective, but it is useful and so I present its

\section{Core Approach}

In order to stand toe-to-toe with Awk, Jacinda has some special facilities to work with streams filtered via regular expressions.

To define a stream from a pattern: {\tt \{\%/Bloom/\}\{`0\}}. {\tt`0} is a line expression; we could write

\section{Type System}

A signal defect of the type system is that it fails to distinguish expressions in the context of a line from expressions in general. In fact expressions in the context of a line form a monad but are not treated as such because explicit typing would be too burdensome and I did not know how to implement implicit coercions.

Consider:

\begin{verbatim}
[y]|> {|`0~/^$/}
\end{verbatim}

This evaluates whether the last line is blank.

\newcommand\Tilde{\char`\~}
\newcommand\Caret{\char`\^}

In a more principled world, we would have {\tt `0 : Ctx Str} and we would force {\tt \{| ... \}} to take an expression of type {\tt Ctx a}. Then we could do something like {\tt \{|(\Tilde/\Caret\$/)"`0\}}... of course this is more prolix.

\subsection{Implementation}

The implementation uses constraint-based typing, as in my Kempe compiler \cite{kempe}. 

Due to various defects in my implementation, polymorphic values can only be instantiated with one type;

\begin{verbatim}
let val sum := [(+)|0 x]
  in sum {|sum (let val l := splitc `0 ' ' in l:i end)} end
\end{verbatim}

is not accepted because {\tt sum} has type {\tt List Int -> Int} at one site and {\tt Stream Int -> Int} at another.

\subsection{Static Typing}

Static typing catches errors before runtime, however Jacinda programs are short, one-off and so the benefit is small.

% editor feedback

\subsubsection{Types as Documentation}

The manpages specify type signatures, e.g.

\begin{verbatim}
\. Binary operator: prior
        (a -> a -> b) -> Stream a -> Stream b
\end{verbatim}

This is more concise and less ambiguous than Awk documentation.

\subsection{Typeclasses}

Type classes are used to witness implementations \cite{adhoc}. Type assignment information is used in the interpreter;

\section{Rewriting Folds}

Often, it is most sensible to express programs in terms of several folds, e.g.

\begin{verbatim}
let
  val tot := (+)|0.0 $1:
  val n := (+)|0.0 [:1.0"$0
in tot%n end
\end{verbatim}

However, it is more efficient if we compute this

\bibliographystyle{plain}
\bibliography{foundations.bib}

\end{document}
